---
title: "Assignment 5"
---

## 1. Using Sample Datasets, create charts using only R graphics functions

# Histogram

```{r}
hist(cars$speed, 
     breaks = 10, 
     col = "skyblue", 
     main = "Histogram of Car Speeds", 
     xlab = "Speed (mph)", 
     ylab = "Frequency",
     border = "white")
```

# Bar Chart(Vertical)

```{r}
# Frequency table for 'speed'
speed_freq <- table(cars$speed)

# Vertical bar chart
barplot(speed_freq, 
        col = "coral", 
        main = "Frequency of Car Speeds", 
        xlab = "Speed (mph)", 
        ylab = "Frequency",
        border = "white")

```

# Bar Chart(Horizontal)

```{r}
barplot(speed_freq, 
        col = "lightgreen", 
        main = "Frequency of Car Speeds (Horizontal)", 
        xlab = "Frequency", 
        ylab = "Speed (mph)",
        horiz = TRUE,
        border = "white")
```

# Pie Chart

```{r}
cyl_freq <- table(mtcars$cyl)  # Create a frequency table for 'cyl'
pie(cyl_freq,  
    col = c("red", "green", "blue", "yellow"),  
    main = "Distribution of Cylinders")

```

# Box Plot

```{r}
# Create a boxplot of mpg by number of cylinders in mtcars dataset
boxplot(mpg ~ cyl, 
        data = mtcars,  
        col = c("red", "blue", "yellow"),  
        main = "MPG by Number of Cylinders",  
        xlab = "Cylinders",  
        ylab = "Miles per Gallon") 

```

# Scatterplot

```{r}
plot(cars$speed, cars$dist, 
     pch = 19, 
     col = "darkorange", 
     main = "Scatterplot of Speed vs. Stopping Distance", 
     xlab = "Speed (mph)", 
     ylab = "Stopping Distance (ft)")
abline(lm(cars$dist ~ cars$speed), col = "blue", lwd = 2)
```

## 1. Using Sample Datasets, create charts using ggplot2 functions

# Histogram

```{r}
library(ggplot2)
ggplot(cars, aes(x = speed)) +
  geom_histogram(bins = 10, fill = "skyblue", color = "white") +  
  labs(title = "Histogram of Car Speeds", 
       x = "Speed (mph)", 
       y = "Frequency") + 
  theme_minimal() 

```

# Bar Chart(Vertical)

```{r}
library(ggplot2)
speed_freq <- as.data.frame(table(cars$speed))
colnames(speed_freq) <- c("Speed", "Frequency")

ggplot(speed_freq, aes(x = Speed, y = Frequency)) +
  geom_bar(stat = "identity", fill = "coral", color = "white") +  
  labs(title = "Frequency of Car Speeds", 
       x = "Speed (mph)", 
       y = "Frequency") +  # Custom labels
  theme_minimal() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  

```

# Bar Chart(Horizontal)

```{r}
library(ggplot2)
speed_freq <- as.data.frame(table(cars$speed))
colnames(speed_freq) <- c("Speed", "Frequency")
ggplot(speed_freq, aes(x = Speed, y = Frequency)) +  
  geom_bar(stat = "identity", fill = "lightgreen", color = "white") +  
  labs(title = "Frequency of Car Speeds (Horizontal)", 
       x = "Speed (mph)", 
       y = "Frequency") +  
  coord_flip() +  
  theme_minimal()  
```

# Pie Chart

```{r}
library(ggplot2)
cyl_freq <- as.data.frame(table(mtcars$cyl)) 
colnames(cyl_freq) <- c("Cylinders", "Frequency")
ggplot(cyl_freq, aes(x = "", y = Frequency, fill = factor(Cylinders))) + 
  geom_bar(stat = "identity", width = 1) +  
  coord_polar(theta = "y") +  
  labs(title = "Distribution of Cylinders", 
       fill = "Cylinders") +  
  scale_fill_manual(values = c("red", "green", "blue", "yellow")) +  
  theme_void()  

```

# Box Plot

```{r}
library(ggplot2)
ggplot(mtcars, aes(x = factor(cyl), y = mpg, fill = factor(cyl))) + 
  geom_boxplot() + 
  scale_fill_manual(values = c("red", "blue", "yellow")) +  
  labs(title = "MPG by Number of Cylinders", 
       x = "Cylinders", 
       y = "Miles per Gallon") +  
  theme_minimal()  

```

# Scatterplot

```{r}
library(ggplot2)
ggplot(cars, aes(x = speed, y = dist)) +
  geom_point(color = "darkorange", shape = 19) +  
  geom_smooth(method = "lm", color = "blue", linewidth = 1.5, se = FALSE) +  
  labs(title = "Scatterplot of Speed vs. Stopping Distance", 
       x = "Speed (mph)", 
       y = "Stopping Distance (ft)") +  
  theme_minimal()  


```

## 3. Export the charts using different formats:

# .PDF

```{r}

pdf("histogram_mpg.pdf") 

# Create the histogram
library(ggplot2)
ggplot(cars, aes(x = speed)) +
  geom_histogram(bins = 10, fill = "skyblue", color = "white") +  
  labs(title = "Histogram of Car Speeds", 
       x = "Speed (mph)", 
       y = "Frequency") + 
  theme_minimal()  

dev.off() 

```

# .JPG

```{r}
library(ggplot2)
speed_freq <- as.data.frame(table(cars$speed))
colnames(speed_freq) <- c("Speed", "Frequency")
jpeg("C:/Users/Ashu/Documents/Data Visualization/AshutoshKadam/vertical_bar_speed.jpg", width = 800, height = 600)
print(
  ggplot(speed_freq, aes(x = Speed, y = Frequency)) +
    geom_bar(stat = "identity", fill = "coral", color = "white") +  
    labs(title = "Frequency of Car Speeds", 
         x = "Speed (mph)", 
         y = "Frequency") +  # Custom labels
    theme_minimal() +  
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
)
dev.off()

```

# .SVG

```{r}
library(ggplot2)
speed_freq <- as.data.frame(table(cars$speed))
colnames(speed_freq) <- c("Speed", "Frequency")
svg("C:/Users/Ashu/Documents/Data Visualization/AshutoshKadam/horizontal_bar_speed.svg", width = 800, height = 600)
print(
  ggplot(speed_freq, aes(x = Speed, y = Frequency)) +  
  geom_bar(stat = "identity", fill = "lightgreen", color = "white") +  
  labs(title = "Frequency of Car Speeds (Horizontal)", 
       x = "Speed (mph)", 
       y = "Frequency") +  
  coord_flip() +  
  theme_minimal()
)
dev.off()
```

# .TIFF

```{r}
library(ggplot2)
cyl_freq <- as.data.frame(table(mtcars$cyl)) 
colnames(cyl_freq) <- c("Cylinders", "Frequency")
tiff("C:/Users/Ashu/Documents/Data Visualization/AshutoshKadam/pie_chart.tiff", width = 800, height = 600)
print(
  ggplot(cyl_freq, aes(x = "", y = Frequency, fill = factor(Cylinders))) + 
  geom_bar(stat = "identity", width = 1) +  
  coord_polar(theta = "y") +  
  labs(title = "Distribution of Cylinders", 
       fill = "Cylinders") +  
  scale_fill_manual(values = c("red", "green", "blue", "yellow")) +  
  theme_void()  
 )
dev.off()
```

# .BMP

```{r}
library(ggplot2)
bmp("C:/Users/Ashu/Documents/Data Visualization/AshutoshKadam/box_plot.bmp", width = 800, height = 600)
print(
  ggplot(mtcars, aes(x = factor(cyl), y = mpg, fill = factor(cyl))) + 
  geom_boxplot() + 
  scale_fill_manual(values = c("red", "blue", "yellow")) +  
  labs(title = "MPG by Number of Cylinders", 
       x = "Cylinders", 
       y = "Miles per Gallon") +  
  theme_minimal()
)
dev.off()
```

# .PNG

```{r}
library(ggplot2)
png("C:/Users/Ashu/Documents/Data Visualization/AshutoshKadam/scatter_plot.png", width = 800, height = 600)
print(
  ggplot(cars, aes(x = speed, y = dist)) +
  geom_point(color = "darkorange", shape = 19) +  
  geom_smooth(method = "lm", color = "blue", linewidth = 1.5, se = FALSE) +  
  labs(title = "Scatterplot of Speed vs. Stopping Distance", 
       x = "Speed (mph)", 
       y = "Stopping Distance (ft)") +  
  theme_minimal()
)
dev.off()
```
