---
title: "Assignment 3"
---

# 1. A Rerun of murrell01.R Boxplot with documentation

```{r}
# Adjusting the margins for the plot: c(bottom, left, top, right)
par(mar=c(3, 4.1, 2, 0))  

# First boxplot for Vitamin C (VC) supplement
boxplot(len ~ dose, data = ToothGrowth,
        boxwex = 0.25,        # Width of the boxes, set to 0.25
        at = 1:3 - 0.2,       # Positioning of the boxes slightly to the left
        subset = supp == "VC",# Data filtered for Vitamin C (VC) supplement
        col = "white",        # Box color set to white
        xlab = "",            # No x-axis label for now
        ylab = "tooth length",# Y-axis label indicating the measured outcome
        ylim = c(0,35))       # Setting the y-axis range from 0 to 35

# Adding x-axis label for Vitamin C doses
mtext("Vitamin C dose (mg)", side = 1, line = 2.5, cex = 0.8)

# Second boxplot for Orange juice (OJ) supplement, added to the same plot
boxplot(len ~ dose, data = ToothGrowth, 
        add = TRUE,           # This ensures the new plot is added on top of the existing one
        boxwex = 0.25,        # Same box width for comparison
        at = 1:3 + 0.2,       # Positioning the boxes slightly to the right
        subset = supp == "OJ")# Data filtered for Orange Juice (OJ) supplement

# Adding a legend to distinguish between the two supplements
legend(1.5, 9,               # Positioning the legend at (1.5, 9) on the plot
       c("Ascorbic acid", "Orange juice"), # Labels for the two supplement types
       fill = c("white", "gray"),          # Corresponding colors: VC is white, OJ is gray
       bty = "n")             # No border around the legend

# Reset margins to the default after the plot
par(mar=c(5.1, 4.1, 4.1, 2.1))

```

# 2. A Rerun of anscombe01.R

## Comparision of regression models:

Anscombe’s quartet consists of four datasets that have nearly identical statistical properties, but they have different distributions and visual patterns.

#### Summary of the Linear Models:

-   **Model 1 (lm1: y1 \~ x1)**:

    -   Coefficients:

        -   Intercept: 3.00

        -   Slope: 0.500

    -   R2: High value, suggests good fit.

    -   Residuals: The residuals are quite small and evenly spread, indicating a reasonably good model.

-   **Model 2 (lm2: y2 \~ x2)**:

    -   Coefficients:

        -   Intercept: 3.00

        -   Slope: 0.500

    -   R2: Similar to model 1, suggests a good fit.

    -   Residuals: Small residuals, but the fit might not be perfect upon visual inspection.

-   **Model 3 (lm3: y3 \~ x3)**:

    -   Coefficients:

        -   Intercept: 3.00

        -   Slope: 0.500

    -   R2: Similar to the previous models.

    -   Residuals: Small, but when visualized, there may be patterns suggesting potential issues (e.g., curvature).

-   **Model 4 (lm4: y4 \~ x4)**:

    -   Coefficients:

        -   Intercept: 3.00

        -   Slope: 0.500

    -   R2: Similar to the others.

    -   Residuals: Larger compared to the others; the single outlier significantly affects the model.

### Key Observations:

-   All four models have nearly identical regression equations, yet their underlying data distributions vary.

-   From the visual inspection of plots, it becomes clear that models 1 and 3 fit reasonably well, but models 2 and 4 require more attention due to apparent non-linearity and an influential outlier, respectively.

-   The linear regression summary alone doesn't reveal these issues, emphasizing the importance of visualization.

## Comparison of different ways to create the plots:

#### Overall, the fancy version enhances visualization with improved aesthetics and standardization, making it more suited for presentation and analysis. The simple version, works well to give a quick overview and for superficial analysis however it is likely to miss finer details. Simple Version:

-   **Color and Line Type**:

    -   Basic colors with default settings.

    -   Uses `plot()` and `abline()` for linear fit, showing the relationship between the variables and the regression line.

    -   Simplicity is good for quick analysis, but it lacks customizability and flair in presentation.

#### Fancy Version:

-   **Enhanced Color and Line Types**:

    -   Uses red points with orange fill (`col = "red", pch = 21, bg = "orange"`), creating a more visually appealing chart.

    -   Regression lines are colored blue for contrast (`abline(mods[[i]], col = "blue")`).

-   **Customization**:

    -   The `xlim` and `ylim` are set to standardize the axis ranges across all plots, making comparisons between the datasets easier.

    -   The layout (`par(mfrow = c(2, 2)`) creates a 2x2 grid to display all four plots together, which is very useful for comparing the datasets side by side.

-   **Labels and Titles**:

    -   An overarching title (`mtext()`) adds context to the entire figure.

    -   Margins (`mar` and `oma`) are adjusted for better spacing.

# 3. Fine Tuning Charts

a\.

```{r}
plot(anscombe$x1,anscombe$y1, family="serif")

```

b\.

```{r}
plot(anscombe$x1,anscombe$y1, col="red")

```

c\.

```{r}
plot(anscombe$x1,anscombe$y1, pch=24)
```
